The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-04-25 paper/ms.tex
Move deltax to appendix, add discussion of epi

%There is a slight complication in the formula for computing $\Delta J_R/J_R$
%and $\Delta J_{\phi}/J_{\phi}$. 
We will later consider the effect on actions of an inaccuracy in the distance from the Sun to the Galactic center, which introduces an offset in the $x$ coordinate, $\Delta x$, of each star when converting to a Galactocentric coordinate system. An offset $\Delta x$ results in an erroneous radius $R_{\text{err}}$ related by
the formula,
\beq
(x+\Delta x)^2 + y^2 = R_{\text{err}}^2\text{.}
\eeq
Keeping only terms to first order in $\Delta x$, we have that,
\beq
\begin{split}
R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
\end{split}
\eeq
Averaging over the circle, we therefore have that,
\beq
\avg{\Delta R} = \frac{2}{\pi} \Delta x\text{.}
\eeq
Therefore, the error in $J_R$ and $J_{\phi}$ are related to $\Delta x$ by,
\beq\label{eq:Jr_epi_dx}
\begin{split}
\frac{\Delta J_R}{J_R} &= \frac{4 \Delta x}{\pi A_R} \\
\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{2 \Delta x}{\pi R}\text{.}
\end{split}
\eeq

@@

@@

\appendix
\section{Orbits} \label{app:orbits}
We plot the three orbits considered throughout the work
(Table~\ref{tab:orbits}) in Figure~\ref{fig:plot_orbits}.

>>>>

%There is a slight complication in the formula for computing $\Delta J_R/J_R$
%and $\Delta J_{\phi}/J_{\phi}$.

@@

In each panel of Figure~\ref{fig:many_orbit_wrong_ref}, where relevant, we
include the estimation of the action errors derived under the epicyclic
approximation in Section~\ref{ssec:epi_action}. Note that we consider an error
in the $x$-coordinate $\Delta x$, which is not quite the same as $\Delta R$.
For observations of stars close to us, we have that $\Delta x \sim \Delta R$,
but for the experiment performed in this Section we consider observations of
the star throughout its entire orbit. This introduces a factor of $2/\pi$ when
converting from $\Delta R$ to $\Delta x$, which we derive in
Appendix~\ref{app:deltax}.

The epicyclic approximation is a good predictor of $J_z$, even for halo
orbits. It performs similarly for $J_R$, now underpredicting for halo orbits
and slightly overpredicting for thin-disk orbits. Note that for the particular
orbits we chose, the thin-disk orbit has slightly larger $A_R$ than the
thick-disk orbit, and so we expect the epicyclic approximation to perform
slightly worse. The epicyclic approximation underpredicts $J_{\phi}$ for all
orbits.

@@

\appendix
\section{$\Delta R$-$\Delta x$ Relation} \label{app:deltax}
We considered the effect on actions of an inaccuracy in the distance from the
Sun to the Galactic center, which introduces an offset in the $x$ coordinate,
$\Delta x$, of each star when converting to a Galactocentric coordinate
system. In observations of nearby stars, we have that $\Delta x \sim \Delta
R$. However, for the experiment we performed in Section~\ref{ssec:quant} we
considered observations of a star throughout its entire orbit. Therefore, we
must average $\Delta R$ over the course of the orbit. We derive this relation
now.

An offset $\Delta x$ results in an erroneous radius $R_{\text{err}}$ related by
the formula,
\beq
(x+\Delta x)^2 + y^2 = R_{\text{err}}^2\text{.}
\eeq
Keeping only terms to first order in $\Delta x$, we have that,
\beq
\begin{split}
R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
\end{split}
\eeq
Averaging over the circle, we therefore have that,
\beq
\avg{\Delta R} = \frac{2}{\pi} \Delta x\text{.}
\eeq

\section{Orbits} \label{app:orbits}
We plot the three orbits considered throughout the work
(Table~\ref{tab:orbits}) in Figure~\ref{fig:plot_orbits}.
